---
title: 'Task 4_3: Classification Challenge'
author: "Thomas Ferreira"
date: "16/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
library(data.table)
library(caret)
library(randomForest)
library(rgdal)
```


## Methodology

Without the use of external data, this becomes a supervised classification problem. If external data could be used, the easiest would be be to overlay small area statistics containing the classifaction from the 2011 Census over the Hex map. 

1. I manually classify some areas into one of the three categories.
2. I divide the classfied data into a training and test set. 
3. I run a random forest to model the classification. Random forests have the ability to handle categorical variables in a classification problem. 
4. I evaluate the prediction on the test data using Confusion Matrices.
5. I predict the classification of all Hex's in the full dataset. 



## Data Setup

```{r}

#Download file if it does not exist. 

if (file.exists("sr_hex.csv")==FALSE) {
  download.file(url="https://cct-ds-code-challenge-input-data.s3.af-south-1.amazonaws.com/sr_hex.csv.gz", destfile ="sr_hex.csv.gz", method='curl')
  untar("sr_hex.csv.gz")
  file.remove("sr_hex.csv.gz")
}

#Read in the data
sr <- fread("sr_hex.csv")
#Subset to Water and Sanitation. 
data <- sr[department=="Water and Sanitation" & h3_level8_index!=0,]
```

##Supervised Classification

In this case, I use some information from division classes and suburb names to identify some areas as urban, rural or informal. 

Informal is identified from the `Codegroup` variable that contains the categories `SEWER - INFORMAL SETTLEMENTS` and `WATER  - INFORMAL SETTLEMENTS`. 

Rural is classified from the `OfficialSuburbs` variable where the word `FARM` appears in the name. 

Urban is classified using some intuition and knowledge of the city also using the `OfficialSuburbs` variable. 

```{r cars}
#Informal
data[CodeGroup %in% c("SEWER - INFORMAL SETTLEMENTS", "WATER  - INFORMAL SETTLEMENTS"),
     geo_type:= "Informal"]
#Rual
data[grep("FARM", data[,OfficialSuburbs]), 
     geo_type:= "Rural"]
#Urban
data[OfficialSuburbs=="CAPE TOWN CITY CENTRE", 
     geo_type:= "Urban"]
data[OfficialSuburbs=="STRAND", 
     geo_type:= "Urban"]
data[OfficialSuburbs=="BERGVLIET", 
     geo_type:= "Urban"]
```

##Generate Training and Test Datasets

```{r}
#Keep data with geo_type classifications (from the supervised classification above)
data_learn <- data[geo_type!="",]

set.seed(1345)

## Generate a data partition

inTrain <- createDataPartition(
  y = data_learn[,geo_type],
  p = .75,
  list = FALSE
)


## Create the training and testing datasets
training <- data_learn[ inTrain,]
testing  <- data_learn[-inTrain,]

```

## Random Forest Clasifier

I only use a subset of all the variables. With the training and testing datasets, its important that both sets have the same factors in each variable. THis limited selection. With a larger training and testing set this would not be a problem. 

The confusion Matrices produced by the classifier are still fairly accurate given the limited information used with an error rate of around 7.8%. This can be improved on with more time and a larger training/testing dataset. 


```{r}
rf = randomForest(factor(geo_type) ~ .,  
                  ntree = 500,
                  data = training[,.(geo_type, CreationDate, CompletionDate, factor(CodeGroup))],
                  xtest = testing[,.(CreationDate, CompletionDate, factor(CodeGroup))],
                  ytest = factor(testing[,geo_type]),
                  keep.forest=TRUE)
rf
```


## Predicting on the full dataset. 

```{r}
data$geo_type_prediction <- predict(rf, data[,.(geo_type, CreationDate, CompletionDate, factor(CodeGroup))], type="response")
table(data$geo_type_prediction)
```




